Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision

### 2023-04-18 Milestone_presentation_2/Milestone_presentation_2/Untitled.Rmd
rearrange the page
restructure according to feedback from other students

---
title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      countdown: 55000
      beforeInit: "scripts/my_macro.js"
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
---

@@

@@

**Residuals** summarise what is **not captured by the regression model**. Many different residual-based hypothesis tests are available to detect specific model defects.

.pull-right-33p[

Non-linearity: 
  - **F-test**
  - **RESET test** (Ramsey 1969)
]

.pull-right-33p[

Heteroskedasticity: 
  - **White test** (White 1980)
  - **BP test** (Breusch and Pagan 1979)

@@

.pull-right-33p[

Non-normality: 
  - **SW test** (Shapiro and Wilk 1965)
  - **Jarque–Bera test** (Jarque and Bera 1980)

@@

.center[
.caption[
Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
]
]

@@

.center[
.caption[
Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
]
]

Many different hypothesis tests are available to detect specific model defects.

Non-linearity: 
  - **F-test**
  - **RESET test** (Ramsey 1969)

Heteroskedasticity: 
  - **White test** (White 1980)
  - **BP test** (Breusch and Pagan 1979)

Non-normality: 
  - **SW test** (Shapiro and Wilk 1965)
  - **Jarque–Bera test** (Jarque and Bera 1980)

.center[
.caption[
Fig. 1: _One classically good looking residual plot and three residual plots with model defects._
]
]

BP and SW tests reject the residual plots exhibiting structure that **they weren’t designed for** (**type III error**).

Most residual-based tests for particular types of departure are also sensitive to **other types of departures** (Cook and Weisberg 1982).

>>>>

---
title: "Advances in Artificial Intelligence for Data Visualization: Automated Reading of Residual Plots with Computer Vision"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      countdown: 55000
      beforeInit: "scripts/my_macro.js"
      countIncrementalSlides: false
      slideNumberFormat: "<%current%>"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
---

@@

.center[

--

--

--

--

```{r fig.height = 2.5, fig.width = 10}
set.seed(10086)

bind_rows(
  dat %>%
    mutate(type = "(A) Good residuals"),
  mod$set_prm("include_z", TRUE)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(B) Non-linearity"),
  heter_model(b = 64)$
    gen(300, computed = select(dat, x, e)) %>%
    mutate(type = "(C) Heteroskedasticity"),
  heter_model(b = 0, e = rand_lognormal(sigma = 0.5))$
    gen(300, computed = select(dat, x)) %>%
    mutate(type = "(D) Non-normality")
) %>%
  mod$plot(remove_grid_line = TRUE, theme = theme_light(base_size = 12)) +
  facet_wrap(~type, ncol = 4, scales = "free") +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank()) +
  xlab("Fitted values") +
  ylab("Residuals")
```

.center[
.caption[
Fig. 1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
]
]

@@

--

Many different **residual-based hypothesis tests** are available to detect specific model defects.

--

.pull-left-33p[

Non-linearity:

- **F-test**

- **RESET test** (Ramsey 1969)

@@

--

.pull-left-33p[

Heteroskedasticity:

- **White test** (White 1980)

- **BP test** (Breusch and Pagan 1979)

--

.pull-left-33p[

Non-normality:

- **SW test** (Shapiro and Wilk 1965)

- **Jarque–Bera test** (Jarque and Bera 1980)

--

.float-left[
But there are **drawbacks** of using them.

- Most residual-based tests for particular types of departure are also sensitive to **other types of departures**, resulting in a phenomenon known as the "**Type III error**" (Cook and Weisberg 1982).

- Other problems such as **outliers** can trigger the rejection (Cook and Weisberg 1999).

@@

.center[
.caption[
Fig. 1: _Example residual plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0._
]
]

@@
